CYO Project Report - Delivery Estimation System

### 2021-07-18 finalReportPrep.Rmd
stable rmd file

---
title: "CYO Project Report - Delivery Estimation System"
author: "Asham Vohra"
date: "6/21/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

@@

# Overview
Prediction systems play a cruical role in the modern world. With advent of technology and availability of data, these systems are being leveraged by retail chains, banks, colleges among other businesses. Depending on the use case, these systems are used to estimate delivery time, sales figures, carry out inventory management, predict occurrence of a disease, among others.

Inspired by what all problems these systems can solve, we have attempted to build a delivery prediction system based on publically available [dataset of orders made at Olist Store](https://www.kaggle.com/olistbr/brazilian-ecommerce), a Brazilian ecommerce.

The dataset used here is hosted at [Kaggle website](https://www.kaggle.com). The dataset is provided by Olist itself, which is the largest department store in Brazilian marketplaces. Olist connects sellers to the customers. It is the sellers, who are responsible to fulfill any orders placed on the platform using Olist's logistic partners.

@@

So in the geolocation data set, zip code prefix refered to the first part of the CEP and it varied from 1 to 5 digits in the dataset available to us. Because the zip code prefix does not represent the precise value, it was not unique and had many duplicate values with varying value of lat, lng for same zip code prefix.

@@

# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
# intially we can see 
# geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()

@@

# so raw data
raw_data <- orders_dataset %>% 
  inner_join(order_items_dataset,by='order_id') %>%
  inner_join(sellers_dataset,by='seller_id') %>% 
  inner_join(customers_dataset,by='customer_id') %>%  
  inner_join(products_dataset,by='product_id') %>%
  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
         product_id=as.factor(product_id), 
         seller_id=as.factor(seller_id),
         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
         customer_unique_id=as.factor(customer_unique_id),
         customer_zip_code_prefix=as.factor(customer_zip_code_prefix),
         product_category_name=as.factor(product_category_name),
         customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(5),pad=c('0'),side='right')) %>%
  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-order_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
```

@@

Having wrangled the datasets individually we combined them into a single dataset. While the dataset had many attributes, based on their frequency and analysis, we zeroed on the below fields. In addition, it is important to note that for model development we created outcome column number_of_hours based on existing fields i.e. order_delivered_customer_date and order_purchase_timestamp as this is what we would want to predict whenever we get a new order i.e. given this order, how many number of hours are required to deliver the product to the customer.

@@

### Performance Metric

@@

#### Current RMSE

@@

This RMSE i.e. our performance metric was later used as benchmark for modeling. The delivery prediction system being built attempted to improve on this metric.

@@

```{r Method to reduce levels in customer_zip_code_prefix, include = FALSE}
# trimming zip code to three digits to be able to extract zip code component which represents region, sub-region and sector data.
compute_rough_customer_location <- function(df) {
  df %>% mutate(customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(3),pad=c('0'),side='right')) %>% 
    mutate(customer_loc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) 
  #%>% select(-customer_zip_code_prefix)
}
```

@@

## Paritioning dataset for training and validation

Before using the wrangled data for analysis and data modeling, we partitioned the data into training and validation data sets. The idea was to keep validation data set aside and use only training data for any analysis, model development and tuning.

@@

The data sets were named as below for furture reference:

@@

For analysis, we used only the training dataset.

Since we needed to build delivery prediction system which could estimate number of hours required for delivery for new orders, while maintaining minimum RMSE, we used our intution and understanding of the domain to guide us in deriving insights from available predictors. The same are shared below:

* _**Distance between seller and customer**_ could likely influence delivery estimate.
* _**Delay in payment approval**_ could influence delivery time as it would delay shippment from the seller. This could be related to customer zip code as often some customer locations require fraud analysis before payment is approved.
* _**Delay in handover of product**_ from seller to carrier would influence delivery time as any time spent there is wasted. This could be seller related as some sellers maybe known to take more time in processing order compared to other sellers. 
* Some _**customer locations**_ maybe difficult to reach or need special delivery mechanism due to their remote nature. Such locations may need more delivery time compared to other locations irrespective of the distance involved. We need not analyse for each customer location but a rough area could be used based on combination of zip codes.
* _**Weight of the product**_ could influence delivery time as heavy items say bed, furniture may be transported slowly while lighter items maybe shipped more quickly.
* _**Price of the product**_ could influence delivery time. Expensive items may require considerable more attention and could influence delivery like priority shipping or longer delivery time due to caution.
* _**Volume of the product**_ could influence delivery time as smaller products can be shipped together while others may need separate truck and therefore more time for dedicated truck to be available.
* _**Product category**_ could be used for analysis as different product categories would have different packing, assembly, processing requiremts and shelf life would could lead to impact in delivery time. 
* _**Freight value**_ could indicate complexity of shipping like fragility of the product, distance, toll booth charges on the way among others. This could be considered for analysis.
* _**Customer state**_ could influence the delivery estimate due to state border checks, paper work before goods delivery among others.
* Different _**sellers**_ may take different amount of time or sell specific type of products which may require different shipping time. So products from a seller selling apparels may be delivered quickly while products from one selling customised products or say furniture may take relatively more time. This could be useful information for predicting delivery time.

To start building our algorithm, we further paritioned our training data set i.e. training_dataset into a split of 80/20 between train_subset and test_subset to ensure we use only the train_subset for training and tuning the model and test_subset to evaluate the model and measure RMSE. This would ensure there is no data leak. The split like earlier was done such that distribution of outcome i.e. number_of_hours was similar in both training and test data

TODO
Use target encoding for categorical variables
Use preprocess for scaling data against training data set. Then use mean, sd from this trained data for test data.

Derived data set should be generated later on observations
- geolocation usage to capture distance
- delay and mean delay in handling to carrier
- number of hours for delivery should be generated later

We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
```{r Split data for training, include=FALSE}
# Let's split training_dataset into two parts
set.seed(1,sample.kind = 'Rounding')
training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% compute_rough_customer_location()

@@

==TODO note penalty data should be calculated again if train_subset is changed. Only trained data should be used for test, so that there is no data leak.

@@

### Product category effects
As part of this section, we looked into our intution that different product categories may have different delivery timelines and explored its influence by incoporating it in our model for analysis.

Since product categories are categorical in nature, we applied target encoding on the product categories attribute.

===TODO talk about weight
===TODO talk about need of regularization.

Below histogram depicts how mean rating of movie across movies is distributed.

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

@@

@@

  weights <- seq(0,6,3)

@@

@@

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

@@

weight_to_use <- mean(weight_to_use_per_partition) 
# 10

@@

@@

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

@@

penalty_data is computed on train_subset and is used with test_subset as historical learning of efficiency of each seller in delivering to carrier
Similarly delay in payment approval is used to incorporate role of customer zip code in delaying this processing, which maybe due to fraud analysis etc.
```{r Prepare training and test data for model tuning and development,include=FALSE}
dataset_list <- prepare_data_for_training_testing(train_subset,test_subset,FALSE)
train_subset <- dataset_list[[1]]
test_subset <- dataset_list[[2]]

@@

5 fold cross validation
80/20 split
TODO remove p=0.8 from knn trainControl as it is not used.

>>>>

---
title: "CYO Project Report - Delivery Estimation System"
author: "Asham Vohra"
date: "6/21/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
---

@@

# Overview
Prediction systems play a crucial role in the modern world. With advent of technology and availability of data, these systems are being leveraged by retail chains, banks, colleges among other businesses. Depending on the use case, these systems are used to estimate delivery time, sales figures, carry out inventory management, predict occurrence of a disease etc.

Inspired by what all problems these systems can solve, we have attempted to build a delivery prediction system based on publicaly available [dataset of orders made at Olist Store](https://www.kaggle.com/olistbr/brazilian-ecommerce), a Brazilian e-commerce.

The dataset used here is hosted at [Kaggle website](https://www.kaggle.com). The dataset is provided by Olist itself, which is the largest department store in Brazilian marketplaces. Olist connects sellers to the customers. It is the sellers, who are responsible to fulfil any orders placed on the platform using Olist's logistic partners.

@@

So in the geolocation data set, zip code prefix referred to the first part of the CEP and it varied from 1 to 5 digits in the dataset available to us. Because the zip code prefix does not represent the precise value, it was not unique and had many duplicate values with varying value of lat, lng for same zip code prefix.

@@

# Important: many geolocation_city have same name but different ascent component. Example: niteroi and niterói
# initially we can see 
# geolocation_dataset %>% dplyr::count(geolocation_zip_code_prefix,geolocation_city,geolocation_state) %>% nrow()

@@

# so raw data
raw_data <- orders_dataset %>% 
  inner_join(order_items_dataset,by='order_id') %>%
  inner_join(sellers_dataset,by='seller_id') %>% 
  inner_join(customers_dataset,by='customer_id') %>%  
  inner_join(products_dataset,by='product_id') %>%
  mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),
         product_id=as.factor(product_id), 
         seller_id=as.factor(seller_id),
         seller_zip_code_prefix=as.factor(seller_zip_code_prefix),
         customer_unique_id=as.factor(customer_unique_id),
         product_category_name=as.factor(product_category_name),
         customer_zip_code_prefix=str_pad(customer_zip_code_prefix, c(5),pad=c('0'),side='left')) %>% 
  select(-customer_id,-order_status,-seller_city,-seller_state,-customer_city,-order_id,-product_id,-seller_zip_code_prefix,-customer_unique_id)
```

@@

Having wrangled the datasets individually we combined them into a single dataset. While the dataset had many attributes, based on their frequency and our intuition shared earlier we zeroed on the below fields for further consideration. In addition, it is important to note that for model development we created outcome column number_of_hours based on existing fields i.e. order_delivered_customer_date and order_purchase_timestamp as this is what we would want to predict whenever we get a new order i.e. given this order, how many number of hours are required to deliver the product to the customer.

@@

## Performance Metric

@@

### Current RMSE

@@

This RMSE i.e. our performance metric was later used as benchmark for modelling. The delivery prediction system being built attempted to improve on this metric.

@@

```{r Method to reduce levels in customer_zip_code_prefix, include = FALSE}
# trimming zip code to three digits to be able to extract zip code component which represents region, sub-region and sector data.
compute_rough_customer_location <- function(df) {
  df %>% mutate(customer_loc = as.numeric(str_extract(customer_zip_code_prefix,"\\d\\d\\d"))) 
}
```

@@

## Partitioning dataset for training and validation

Before using the wrangled data for analysis and data modelling, we partitioned the data into training and validation data sets. The idea was to keep validation data set aside and use only training data for any analysis, model development and tuning.

@@

The data sets were named as below for future reference:

@@

```{r Split data for training, include=FALSE}
# We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model, identify weights for target encoding of categorical variables etc.
# Let's split training_dataset into two parts
set.seed(1,sample.kind = 'Rounding')
training_dataset <- training_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% compute_rough_customer_location()

@@

@@

## Feature Analysis and Selection
Since we needed to build delivery prediction system which could estimate number of hours required for delivery for new orders, while maintaining minimum RMSE, we used our intuition and understanding of the domain to guide us in deriving insights from available predictors. The same are shared below:

* _**Distance between seller and customer**_ could likely influence delivery estimate.
* _**Delay in payment approval**_ could influence delivery time as it would delay shipment from the seller. This could be related to customer zip code as often some customer locations require fraud analysis before payment is approved.
* _**Delay in handover of product from seller to carrier**_ would influence delivery time as any time spent there is wasted. This could be seller related as some sellers maybe known to take more time in processing order compared to other sellers. 
* Some _**customer locations**_ maybe difficult to reach or need special delivery mechanism due to their remote nature. Such locations may need more delivery time compared to other locations irrespective of the distance involved. We need not analyse for each customer location but a rough area could be used based on combination of zip codes.
* _**Weight of the product**_ could influence delivery time as heavy items say bed, furniture may be transported slowly while lighter items maybe shipped more quickly.
* _**Price of the product**_ could influence delivery time. Expensive items may require considerable more attention and could influence delivery like priority shipping or longer delivery time due to caution.
* _**Volume of the product**_ could influence delivery time as smaller products can be shipped together while others may need separate truck and therefore more time for dedicated truck to be available.
* _**Product category**_ could be used for analysis as different product categories would have different packing, assembly, processing requirements and shelf life would could lead to impact in delivery time. 
* _**Freight value**_ could indicate complexity of shipping like fragility of the product, distance, toll booth charges on the way among others. This could be considered for analysis.
* _**Customer state**_ could influence the delivery estimate due to state border checks, paper work before goods delivery among others.
* Different _**sellers**_ may take different amount of time or sell specific type of products which may require different shipping time. So products from a seller selling apparels may be delivered quickly while products from one selling customised products or say furniture may take relatively more time. This could be useful information for predicting delivery time.

For analysis, we used only the training_dataset or subset of it, ensuring validation_dataset was left untouched.

### Derived features

#### Distance between seller and customer feature
As part of the dataset we had customer and seller location details available in the form of latitude(lat), longitude(lng). We used this crucial knowledge of their location to determine shortest distance between the two on an ellipsoid.

We decided to look more into this intuition. For this we plotted numbers of hours taken for delivery against the distance between seller and the customer. In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.

```{r Explore distance_between_two_points relation with number_of_hours, echo=FALSE, message=FALSE}
training_dataset  %>% ggplot(aes(distance_between_two_points,number_of_hours)) + geom_point() + geom_smooth() + xlab('Distance between seller and the customer') + ylab('Number of hours taken for delivery')
```
In above plot, we noted that number of hours taken for delivery showed stronger correlation for smaller distances. However, once delivery distance increased beyond a threshold, the correlation weakened between the fields. Though, we could see broader confidence interval for larger distances.

```{r correlation analysis for distance_between_two_points with number of hours, echo=FALSE}
# correlation analysis for distance_between_two_points with number of hours

analysis_output_rows <- sapply(seq(0,2000,1000), function(offset) {  
  list(delay_limit=str_c('>=',offset,' & <',offset+1000),
       correlation=cor(training_dataset %>% 
                         filter(distance_between_two_points>=offset & distance_between_two_points< offset+1000) %>% 
                         pull(distance_between_two_points), 
                       training_dataset %>% 
                         filter(distance_between_two_points>=offset & distance_between_two_points< offset+1000) %>% 
                         pull(number_of_hours))) 
  },simplify=TRUE)

analysis_output_rows %>% t() %>% rbind( c(str_c('>=',3000),
                          round(cor(training_dataset %>% 
                                      filter(distance_between_two_points>=3000) %>%
                                      pull(distance_between_two_points), 
                                    training_dataset %>% 
                                      filter(distance_between_two_points>=3000) %>%
                                      pull(number_of_hours)),7))
                        ) %>% as.data.frame() %>% knitr::kable(caption='Correlation analysis between distance between seller & carrier and number of hours taken for delivery')

rm(analysis_output_rows)
```
This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.

#### Delay in handover of product from seller to carrier feature
As part of the dataset, there were two important attributes i.e. handed_over_to_carrier_date and shipping_limit_date. Here, handed_over_to_carrier_date is the actual timestamp when product ordered was handed over the carrier for delivery. While, shipping_limit_date is the date by which seller should handover the order to the carrier.

Together these fields were considered useful as they could be used to understand if there was a delay or not in handing over product ordered by seller to the carrier. Also, this delay or early handover, could influence delivery timeline even if everything else was to take the same time.

We decided to look more into this intuition. For this, the feature i.e mean_delay_in_handover was computed for each seller. The computed data was then joined with training_dataset and the below graph was used to visualize the relation between delay in handover to carrier and number of hours taken for delivery. In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.

```{r Explore delay_in_handover_to_carrier relation with number_of_hours, echo=FALSE, message=FALSE}
delay_in_handover_to_carrier_analysis_dataset <- training_dataset %>%
  get_delay_in_handover_to_carrier_per_seller_id()

analysis_joined_data <- training_dataset %>%
  left_join(delay_in_handover_to_carrier_analysis_dataset,by='seller_id')

analysis_joined_data %>% ggplot(aes(mean_delay_in_handover,number_of_hours)) + geom_point() + geom_smooth() + xlab('Delay in handover of product to carrier') + ylab('Number of hours taken for delivery')
```

In above plot, we noted that number of hours taken for delivery shows weaker correlation with  delay in handover by seller to the carrier as long as the ordered product is handed over to carried before the shipping_limit_date. However, post the date, we could see relatively stronger correlation between the fields.

```{r correlation analysis for mean_delay_in_handover with number of hours, echo=FALSE}
# correlation analysis for mean_delay_in_handover with number of hours
data.frame(delay_limit=character(0),correlation=numeric(0)) %>% 
  add_row(delay_limit='<= 0',correlation=cor(analysis_joined_data %>% 
                                               filter(mean_delay_in_handover<= 0) %>% 
                                               pull(mean_delay_in_handover), 
                                             analysis_joined_data %>% 
                                               filter(mean_delay_in_handover<= 0) %>% 
                                               pull(number_of_hours))
          ) %>% 
  add_row(delay_limit='> 0',correlation=cor(analysis_joined_data %>% 
                                              filter(mean_delay_in_handover> 0) %>% 
                                              pull(mean_delay_in_handover), 
                                            analysis_joined_data %>% 
                                              filter(mean_delay_in_handover> 0) %>% 
                                              pull(number_of_hours))) %>% 
  knitr::kable(caption='Correlation analysis between delay in handover to carrier and number of hours taken for delivery')
```

This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model. However, as for new orders this information would not be available at time of placing the order, we were required to compute this feature using training data and apply the captured learnings later on test data.

To be able to do this, as shared earlier, the feature i.e. mean_delay_in_handover was computed for each seller based on training dataset and then applied to test dataset.

```{r cleanup for delay in handover to carrier feature analysis, include=FALSE}
rm(delay_in_handover_to_carrier_analysis_dataset,analysis_joined_data)
```

#### Delay in payment approval feature
In addition to above, as part of the dataset, there were two other important attributes i.e. order_purchase_timestamp and payment_approved_at. Here, payment_approved_at is the actual timestamp when payment is approved for a given order. While, order_purchase_timestamp is the timestamp at which product was ordered.

Together these fields were considered useful as they could be used to understand if there was a delay or not in approving the payment. Also, this delay or early handover, could influence delivery timeline even if everything else was to take the same time.

We decided to look more into this intuition. For this, the feature i.e mean_delay_in_payment_approval was computed for each customer_zip_code_prefix. The computed data was then joined with training_dataset and the below graph was used to visualize the relation between delay in payment approval and number of hours taken for delivery. In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.

```{r Explore delay_in_payment_approval relation with number_of_hours, echo=FALSE, message=FALSE}
delay_in_payment_approval_analysis_dataset <- training_dataset %>%
  get_delay_in_payment_approval_per_customer_zip_code()

analysis_joined_data <- training_dataset %>%
  left_join(delay_in_payment_approval_analysis_dataset,by='customer_zip_code_prefix')

analysis_joined_data %>% ggplot(aes(mean_delay_in_payment_approval,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Delay in payment approval') + ylab('Number of hours taken for delivery [log10(n) scale]')
```
In above plot, we noted that number of hours taken for delivery follow a non linear pattern.

```{r correlation analysis for mean_delay_in_payment_approval with number of hours, echo=FALSE}
# correlation analysis for mean_delay_in_payment_approval with number of hours
analysis_output_rows <- sapply(seq(0,125,25), function(offset) {  
  list(delay_in_approval=str_c('>=',offset,' & <',offset+25),
       correlation=cor(analysis_joined_data %>% 
                         filter(mean_delay_in_payment_approval>=offset & mean_delay_in_payment_approval< offset+25) %>% 
                         pull(mean_delay_in_payment_approval), 
                       analysis_joined_data %>% 
                         filter(mean_delay_in_payment_approval>=offset & mean_delay_in_payment_approval< offset+25) %>% 
                         pull(number_of_hours))) 
  },simplify=TRUE)

analysis_output_rows %>% t() %>% rbind( c(str_c('>=',150),
                          round(cor(analysis_joined_data %>% 
                                      filter(mean_delay_in_payment_approval>=150) %>%
                                      pull(mean_delay_in_payment_approval), 
                                    analysis_joined_data %>% 
                                      filter(mean_delay_in_payment_approval>=150) %>%
                                      pull(number_of_hours)),7))
                        ) %>% as.data.frame() %>% knitr::kable(caption='Correlation analysis between delay in payment approval and number of hours taken for delivery')

rm(analysis_output_rows)
```
This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model. However, as for new orders this information would not be available at time of placing the order, we were required to compute this feature using training data and apply the captured learnings later on test data.

To be able to do this, as shared earlier, the feature i.e. mean_delay_in_payment_approval was computed for each customer_zip_code_prefix based on training dataset and then applied to test dataset.

```{r cleanup for delay in payment approval feature analysis, include=FALSE}
rm(delay_in_payment_approval_analysis_dataset,analysis_joined_data)
```

#### Volume feature
As part of the dataset, there were three attributes about product dimensions i.e. product_length_cm, product_height_cm and product_width_cm. Together these fields were considered useful as they could be used to derive volume of the product.

We looked into the intuition that product volume could influence the delivery timelines as larger products need special delivery mechanism and even dedicated trucks at times for shipment, which may not be readily available.

To understand its influence on the delivery timelines, we plotted numbers of hours taken for delivery against the volume of the product ordered in cubic centimeter($cm^3$). In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.

```{r Explore volume relation with number_of_hours, echo=FALSE, message=FALSE}
training_dataset  %>% ggplot(aes(volume,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Volume of ordered product') + ylab('Number of hours taken for delivery [log10(n) scale]')
```
In above plot, we noted that number of hours taken for delivery vary as the product volume changes. There appeared to be stronger influence on delivery timelines for bulkier products beyond a limit.

```{r correlation analysis for volume with number of hours, echo=FALSE}
# correlation analysis for volume with number of hours

analysis_output_rows <- sapply(seq(0,210000,70000), function(offset) {  
  list(delay_limit=str_c('>=',offset,' & <',offset+70000),
       correlation=cor(training_dataset %>% 
                         filter(volume>=offset & volume< offset+70000) %>% 
                         pull(volume), 
                       training_dataset %>% 
                         filter(volume>=offset & volume< offset+70000) %>% 
                         pull(number_of_hours))) 
  },simplify=TRUE)

analysis_output_rows %>% t() %>% rbind( c(str_c('>=',280000),
                          round(cor(training_dataset %>% 
                                      filter(volume>=280000) %>%
                                      pull(volume), 
                                    training_dataset %>% 
                                      filter(volume>=280000) %>%
                                      pull(number_of_hours)),7))
                        ) %>% as.data.frame() %>% knitr::kable(caption='Correlation analysis between volume and number of hours taken for delivery')

rm(analysis_output_rows)
```

This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.

### Categorical features

#### Product category effects
We decided to look into our intuition that different product categories may have different delivery timelines and explore its influence on the prediction results.

Below errorbar plot depicts how average number of hours taken for delivery vary across product categories while at the same time capturing the confidence interval of number of hours taken for delivery for each product category.

```{r Explore product_category_name relation with number_of_hours, echo=FALSE}
training_dataset %>% group_by(product_category_name) %>%
  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% 
  mutate(product_category_name=reorder(product_category_name,mean)) %>% 
  ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
  geom_point()+ geom_errorbar() + theme(axis.text.x = element_text(angle=90,size=8)) + xlab('Product Categories') + ylab('Avg number of hours')
```

In above plot, we noted that average number of hours taken for delivery vary considerably from one product category to another. So our intuition that product categories influence delivery time was confirmed.

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, product categories are categorical in nature and they are quite a few categories, incorporating them directly would not have worked for many machine learning models which require numerical values. Additionally in some cases, each category could end up being treated as a predictor which is also the case with one-hot encoding.

However, to convert categorical variables to numeric values and still be able to limit features and make model training and evaluation less expensive, we applied **target encoding** on the product categories attribute.

But to do that, first, we checked number of observations per product category and noticed that the distribution could require **regularization** to prevent overfitting. Secondly, target encoding could be erroneous if done against the random dataset.

To solve these, we identified the regularization weight to use by carrying out target encoding for different datasets set of weights, selecting single weight per dataset based on performance metric i.e. RMSE and then we took a mean of weights identified across the datasets.

Once weight was identified, to avoid any randomness due to choice of the dataset, we computed target encoding for product category attribute across 10 datasets and then took a mean for them to settle at the final target encoded values to use for the concerned feature.

For the above analysis, knn model was used, given that it does not make assumptions about the dataset, which was useful in the early phase of analysis. However, to use knn model, we **scaled and normalized** the features as knn model uses Euclidean distance. Scaling and normalization was carried our using training data and then used for test data.

@@

      tuningControl=trainControl(method = "cv", number = 3)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(41,47,53)),trControl=tuningControl)

@@

#### Seller id effects
Next we looked into the intuition that orders for products from different sellers based on what they sell, how they operate could influence delivery timelines.

Below errorbar plot depicts how average number of hours taken for delivery vary across sellers while at the same time capturing the confidence interval of number of hours taken for delivery for each seller. Since there were **`r training_dataset %>% dplyr::distinct(seller_id) %>% nrow()`** sellers, we sampled **300** of them to represent their influence effectively in the graph.

```{r Explore seller_id relation with number_of_hours, echo=FALSE, warning = FALSE}
set.seed(1, sample.kind = 'Rounding')
training_dataset %>% group_by(seller_id) %>%
  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>%
  mutate(seller_id=reorder(seller_id,mean)) %>% 
  filter(seller_id %in% sample(seller_id,300,replace=FALSE)) %>%
  ggplot(aes(x=seller_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
  geom_point()+ geom_errorbar() + theme(axis.text.x = element_blank()) + xlab('Seller Ids') + ylab('Avg number of hours')
```
In above plot, we noted that average number of hours taken for delivery vary considerably from one seller to another. So our intuition that sellers influence delivery time was confirmed.

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, seller ids similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(seller_id) %>% nrow()`** sellers, which are a lot. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for seller ids we applied target encoding and followed strategy used for product categories.

@@

  weights <- seq(0,9,3)

@@

weight_to_use <- mean(weight_to_use_per_partition)  
# 8

@@

#### Customer state effects
Next we looked into the intuition that customer states can have different regulations, paper work, approvals which could influence delivery timelines.

Below errorbar plot depicts how average number of hours taken for delivery vary across customer state while at the same time capturing the confidence interval of number of hours taken for delivery for each customer state.

```{r Explore customer_state relation with number_of_hours, echo=FALSE, warning = FALSE}
set.seed(1, sample.kind = 'Rounding')
training_dataset %>% group_by(customer_state) %>%
  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>%
  mutate(customer_state=reorder(customer_state,mean)) %>% 
  ggplot(aes(x=customer_state,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
  geom_point()+ geom_errorbar() + theme(axis.text.x = element_text(angle=90,size=8)) + xlab('Customer States') + ylab('Avg number of hours')
```
In above plot, we noted that average number of hours taken for delivery vary considerably from one customer state to another. So our intuition that customer state influences delivery time was confirmed.

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, customer states similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_state) %>% nrow()`** customer states. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for customer states we applied target encoding and followed strategy used for product categories.

@@

      tuningControl=trainControl(method = "cv", number = 3)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(34,41,48)),trControl=tuningControl)

@@

@@

#### Customer location(loc) effects
Next we looked into the intuition that rough information about customer location i.e. combination of co-located zip codes can influence the delivery timelines as customer location can be remote or require different means of transport, among others.

Below errorbar plot depicts how average number of hours taken for delivery vary across customer locations while at the same time capturing the confidence interval of number of hours taken for delivery for each customer location.

```{r Explore customer_loc relation with number_of_hours, echo=FALSE, warning = FALSE}
set.seed(1, sample.kind = 'Rounding')
training_dataset %>% group_by(customer_loc) %>%
  summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>%
  mutate(customer_loc=reorder(customer_loc,mean)) %>% 
  ggplot(aes(x=customer_loc,y=mean,ymax=mean+2*se,ymin=mean-2*se)) +
  geom_point()+ geom_errorbar() + theme(axis.text.x = element_blank()) + xlab('Customer Locations') + ylab('Avg number of hours')
```
In above plot, we noted that average number of hours taken for delivery vary considerably from one customer location to another. So our intuition that customer location influences delivery time was confirmed.

Given the influence, we decided to consider this as a feature for further analysis for inclusion in the model. However, customer locations similar to product categories are categorical in nature and there are **`r training_dataset %>% dplyr::distinct(customer_loc) %>% nrow()`** customer locations. Incorporating them as it is would not have worked for many machine learning models which require numerical values. As a result, for customer states we applied target encoding and followed strategy used for product categories.

@@

      tuningControl=trainControl(method = "cv", number = 3)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(35,43,49)),trControl=tuningControl)

@@

### Other features

#### Price feature
We looked into the intuition that order item's price could influence the delivery timelines as expensive items ordered may have priority shipping or be delivered with caution.

To understand its influence on the delivery timelines, we plotted numbers of hours taken for delivery against the price of the ordered item. In order to see the trend, gam method based smoothing curve was plotted on the graph.

```{r Explore price relation with number_of_hours, echo=FALSE, message=FALSE}
training_dataset  %>% ggplot(aes(price,number_of_hours)) + geom_smooth() + xlab('Price of ordered item') + ylab('Number of hours taken for delivery')
```
In above plot, we noted that number of hours taken for delivery vary as the price of the ordered item changes. It appeared to increase significantly as the price of ordered item increased and decreased when the price reached towards higher limits. Though confidence interval indicated higher variability in delivery timelines at higher prices.

This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.

#### Product weight feature
Next we looked into the intuition that product weight can influence the delivery timelines as heavier items like furniture, bed may be transported slowly compared to lighter items.

For this we plotted numbers of hours taken for delivery against the weight of the product ordered in grams. In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.

```{r Explore product weight relation with number_of_hours, echo=FALSE, message=FALSE}
training_dataset  %>% ggplot(aes(product_weight_g,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Weight of ordered product in grams') + ylab('Number of hours taken for delivery [log10(n) scale]')
```
In above plot, we noted that number of hours taken for delivery vary as the product weight changes. This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.

#### Freight value feature
Next we looked into the intuition that freight value can help determine the delivery timelines as higher freight value could indicate complexity in shipment, extra processing which would need time, among others.

For this we plotted numbers of hours taken for delivery against the freight value of the product ordered. In order to see the trend in this over-plotted graph, gam method based smoothing curve was also plotted on the graph.

```{r Explore freight value relation with number_of_hours, echo=FALSE, message=FALSE}
training_dataset  %>% ggplot(aes(freight_value,log10(number_of_hours))) + geom_point() + geom_smooth() + xlab('Freight value of ordered product') + ylab('Number of hours taken for delivery [log10(n) scale]')
```
In above plot, we noted that number of hours taken for delivery vary as the freight value of product changes. It appeared to increase as freight value increased for smaller freight values, then plateaued and decreased as freight value reached higher limits The decrease could be due to priority shipping coming at a cost. This behavior could be useful in estimating delivery timelines. We decided to consider this as a feature for further analysis for inclusion in the model.

## Model Selection
```{r Prepare training and test data for model tuning and development,include=FALSE}
dataset_list <- prepare_data_for_training_testing(train_subset,test_subset,FALSE)
train_subset <- dataset_list[[1]]
test_subset <- dataset_list[[2]]

@@

Post evaluation with multiple features, the final set of features identified for model selection were as below::

```{r Final set of features identified, echo=FALSE}
train_subset %>% colnames() %>% knitr::kable(col.names=c('Feature'),caption='Features List')
```

To start building our model, training and test subset were prepared. For this, we further partitioned our training data set i.e. training_dataset into a split of 80/20 between train_subset and test_subset to ensure we use only the train_subset for training and tuning the model and test_subset to evaluate the model and measure RMSE. This would ensure there is no data leak. The split like earlier was done such that distribution of outcome i.e. number_of_hours was similar in both training and test data.

The derived features selected earlier were prepared based on train_subset and applied over test_subset, to ensure we maintain the environmental condition where test data is not known to us.
